---
title: "Craft Beer"
subtitle: <span style="color:blue">Alcohol Content and Bitterness Preference Analysis</span>
authors: "Michael Catalano + Hayley Horn"
date: "October 16, 2018"
output:
  html_document: default
  pdf document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#Prepared by Michael Catalano + Hayley HOrn
#6306-404 Case Study #1
```

```{r Libraries = FALSE}
#These are the packages used in this project
library(blogdown)
library(data.table)
library(DT)
library(ggplot2)
library(kableExtra)
library(plyr)
library(RColorBrewer)
```

<br><br>
Beer can be made using a range of ingredients to adjust the color, type and flavor profile. This analysis focuses on two measurements of beer composition, and some geographic details. The first measurement is Alcohol By Volume (ABV) which is a percentage that represents how much of a beer is alcohol versus other ingredients. The other measurment we will consider is International Business Units (IBU) which is measured on a scale of 0 to 100 and describes the bitterness from hops in a beer. By law in the United States, the ABV must be listed on each beer to give customers an idea about how much alcohol is in each brew. 

For this analysis, we have taken data from 2,410 craft beers and sales data from 558 US breweries to identify trends or regionality for preference. We use craft brewery data versus national brands as there is more product experimentation and product churn.
<br>

```{r Data_Load, echo=FALSE}
#This section pulls the source data files

Beer <- data.frame(read.csv("https://raw.githubusercontent.com/TheDataCurator/SMU-MSDS/master/Case%20Study/Beers.csv", header = TRUE, sep = ",", dec = ".")) #Load the Beer data semicolon delimitied text file

Brew <- data.frame(read.csv("https://raw.githubusercontent.com/TheDataCurator/SMU-MSDS/master/Case%20Study/Breweries.csv", header = TRUE, sep = ",", dec = ".")) #Load the Brewery data

names(Beer) <- c("Beer_Nm","Beer_ID","ABV","IBU","Brew_ID","Style","Ounces")#Rename for consistency and easier linking
```

<br><span style="color:blue">
<b>There is a great amount of variablility in number of breweries by state.</b><br></span>
Colorado leads the country in breweries, followed by California, Michigan, Oregan and Texas.

```{r Brewery_Count, echo=FALSE}
#How many breweries are in each state

Brew$State = trimws(Brew$State, "l") #Remove Whitespace

Brew_summary <- data.frame(count(Brew$State)) #Create dataframe

names(Brew_summary) <- c("State","Brew_Count") #Create more user friendly labels

#1. How many breweries are present in each state
datatable(Brew_summary, rownames = FALSE, filter = "top", options = list(pageLength = 5, scrollX = T),class = 'cell-border stripe',caption = 'Count of Craft Breweries by State.', width = 500 ) #Creates table for easy filtering and view of breweries by State
```
<br><span style="color:blue">
<b>First we take a look at the first and last 6 records of the beer and brewery data</b></span>
We merged the data from both files for a consolidated view of brewery information, beer information, including the ingredients of interest ABV and IBU and other attributes were pulled in to for some flexibility in analysis. The top of the file looks fairly complete and clean.

The end of the file indicates there are IBU values missing, so next we look to see the extent to which the field is incomplete.
<br>
```{r merge, echo = FALSE}
#2. Merge beer data with the breweries data. Print the first six observations and last six observations to check the merged file.

BandBs <- (merge(Beer,Brew, by = "Brew_ID")) #Merges the data sets 

kable(head(BandBs,6), longtable = TRUE, booktabs=TRUE, caption="The first 6 records in the Beer and Brewery file") %>% kable_styling(full_width = FALSE) # Print the first 6 lines of the Beer merged beer & brewery data

kable(tail(BandBs,6), longtable = TRUE, booktabs = TRUE,caption="The last 6 records in the Beer and Brewery file") %>% kable_styling(full_width = FALSE)# Print the last 6 lines of the merged beer & brewery data
```
<br><span style="color:blue">
<b>Investigate extent of incomplete data in the data sets.</b></span><br>
After looking at the head and tail, we looked to see how many fields were incomplete. The field that was most commonly incomplete was IBU, which was expected, and some ABVs which was not expected as they are required by law to be shared with a consumer. We do however still have enough information to continue with the analysis.
<br>

```{r incompletes, echo = FALSE}
#3. Report the number of NAs in each column
na_BandBs <- sapply(BandBs, function(y) sum(length(which(is.na(y))))) #counts the NAs

#There are 2 fields with NAs, 1005 IBUs, 62 ABV
kable(na_BandBs, longtable = TRUE, booktabs = TRUE,caption="Count of NAs by field in the Beer and Brewery file",escape = TRUE) %>% kable_styling("striped", position = "center")
```
<br><span style="color:blue">
<b>For perspective, we looked at the median alcohol content and bitterness unit for each state.</b></span><br>
There was more variability in IBU by State, versus the ABV which you can see groups of states with a similar ABV%.<br>

```{r ABV_IBU_y_State, echo = FALSE}
#Compute median ABV and IBU by State

ABV_state <- aggregate(BandBs$ABV, list(BandBs$State), median, na.rm = TRUE) #aggregates the ABV data
names(ABV_state) <- c("State", "ABV_Pct") #rename the fields

IBU_state <- aggregate(BandBs$IBU, list(BandBs$State), median, na.rm = TRUE) #aggregates the IBU data
names(IBU_state) <- c("State", "IBUs") #renames the fields

IBU_state <- IBU_state[-42,] #Remove NA record

#Plot the chart to compare - for #4
getPalette = colorRampPalette(brewer.pal(9, "PuBu"))

par(mfrow=c(2,1))
colourCountABV = length(unique(ABV_state$State))
ABV_state$State <- factor(ABV_state$State, levels = ABV_state$State[order(-ABV_state$ABV_Pct)])
ggplot(data = ABV_state , aes(x = ABV_state$State , y = ABV_state$ABV_Pct, fill = ABV_state$State)) + geom_bar(stat = 'identity', show.legend = FALSE) + labs(title = "Beer ABV by State", x = 'State', y = 'ABV' , fill = 'State') + theme_bw(10) +
  theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
  theme(legend.title.align = 0.5) + theme(plot.title = element_text(hjust = 0.5))+
   scale_fill_manual(values = rev(getPalette(colourCountABV)))
  #ABV by State bars into descending order, and graph

colourCountIBU = length(unique(IBU_state$State))
IBU_state$State <- factor(IBU_state$State, levels = IBU_state$State[order(-IBU_state$IBUs)])

ggplot(data = IBU_state , aes(x = IBU_state$State , y = IBU_state$IBUs, fill = IBU_state$State)) + geom_bar(stat = "identity", show.legend = FALSE) + labs(title = "Beer IBUs by State", x = "State", y = "IBUs" , fill = 'State') + theme_bw(10) +
  theme(axis.text.x = element_text(angle = 90,hjust = 1)) +
  theme(legend.title.align = 0.5) + theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = rev(getPalette(colourCountIBU)))
  #IBU by State bars into descending order, and graph
```
<br><span style="color:blue"><b>Here is a closer look at the details for the highest values for ABV and IBU.</b>
<br></span>
There is a tie in ABV, which is in line with what we saw in the chart above, where multiple states shared the same ABV value.<br>
```{r Max ABV, echo=FALSE}
#5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
library("data.table")
ABV_state_sorted <- setorderv(ABV_state, "ABV_Pct")[]
tail(ABV_state_sorted,2)
#There were two states with the highest ABV- DC and KY
IBU_state_sorted <- setorderv(IBU_state, "IBUs")[]
tail(IBU_state_sorted,1)
#The state with the highest IBU is ME
```
<br><b><span style="color:blue">
We wanted to share the range of the alcohol content and bitterness </b><br></span>
You can see from the boxplot that there are outliers- which can affect the analysis, but they are valid data points.
```{r Summary Statistics, echo=FALSE}
#6. Summary Statistics for ABV
summary(ABV_state$ABV_Pct)
boxplot(ABV_state$ABV_Pct, horizontal = TRUE, col = "light blue")

summary(IBU_state$IBUs)
boxplot(IBU_state$IBUs, horizontal = TRUE, col = "bisque")
```

<br><b> <span style="color:blue">We plotted bitterness and alcohol content </b><br></span>
The scatterplot was created to see if there was a relationship between the IBU and ABV at the state level, but the results were unremarkable.

```{r Scatterplot, echo=FALSE}
#7. Is there a relationshop between bitterness of beer and its alcohol content. Use a scatterplot to visualize the data.

ABV_IBU_combined <- (merge(ABV_state,IBU_state, by = "State"))#Re-combine ABV and IBU tables

#Plot the scatterplot
  ggplot(ABV_IBU_combined, aes(x=IBUs, y=ABV_Pct)) + 
  geom_point()+
  geom_smooth(method=lm) +
  ylab("ABV") + labs(title="ABV and IBU Scatter Plot") + theme(plot.title = element_text(hjust = 0.5))
```
<br><b><span style="color:blue">Approximately 44% of the variance of IBUs can be explained by ABV.</b></span><br>
The study showed very little correlation (r = 0.282, p-value = 0.0471) between ABV and IBUs using scatter plot and Pearson's Product correlation formula. 
```{r correlation, echo=FALSE}
# Calculate correlation between ABV and IBUs
cor.test(ABV_IBU_combined$ABV_Pct, ABV_IBU_combined$IBUs)
```
<br><span style="color:blue"><b>In Summary</b></span>

There was very little evidence of a relationship between the ABV and IBU. The state of Maine had the highest IBU content and there was a tie for alcohol content between the District of Columbia and the state of Kentucky.

For further analysis, we recommend applying demographic information to the analysis to determine if different generational segments have preferences, and regionality of preference with the target generational segments.  
