---
title: "6306-404 U5 Live Session Assignment"
author: "Hayley Horn"
date: "September 30, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("C:/Users/ha406272/Dropbox/SMU/")
#setwd("C:/Users/hayle/Dropbox/SMU/")
```

###Purpose
Consumer beer preference is a changing and the ability to understand, adapt and forecst the consumer's changing tastes is paramount for future profibility. We have taken data 2410 craft beers and sales data from 558 US breweries, and we have used that information to find a insights to understand components affecting beer consumption. Consumers are not always able to articulate preferences consistent with industry terms, so this approach removes the burdern of industry knowledge from the consumer- they have voted with their pocketbooks.






####Questions- Do not include specifically
Questions

1.	How many breweries are present in each state?
2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
3.	Report the number of NA's in each column.
4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
6.	Summary statistics for the ABV variable.
7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.




###1.Data Import

```{r Data_Load, echo=TRUE}
#a: Load the semicolon delimitied text file
Beers <- data.frame(read.csv("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_6/beers.csv", header = TRUE, sep = ",", dec = "."))
str(Beers)
Beers

Breweries <- data.frame(read.csv("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_6/breweries.csv", header = TRUE, sep = ",", dec = "."))
str(Breweries)
Breweries

#Give the columns names
names(y2016) <- c("Name","Gender","Count")

#b: Summarize and look at the structure of the file
summary(y2016)
str(y2016)

#c: Find the errant row
display <- grep("yyy",y2016$Name)
display

#d: Remove the duplicate row
y2016 <- y2016[-212,]
```

###2.	Data Merging

```{r Data_Merging_2, echo=TRUE}
#a: Load the comma delimited text file
y2015 <- data.frame(read.delim("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_5/yob2015.txt", header = FALSE, sep = ",", dec = "."))

    #Give the columns names
names(y2015) <- c("Name","Gender","Count")

#b: Display last 10 rows in the data file
tail(y2015, n = 10)


```
```{r Data_Merging_2c, echo=TRUE}
#c: load the library and merge the 2 files
     #####This is not final- need to dedupe the rows#####
final <- (merge(y2016, y2015, by = c("Name","Gender")))
names(final) <- c("Name","Gender","2016_Count", "2015_Count")

```

###3.	Data Summary

```{r Data_Summary_3, echo=TRUE}
#a: Create column named Total with combined totals

final$Total <- c(final$`2016_Count` + final$`2015_Count`)

#b: Sort and select top 10 names
final <- final[order(-final$Total),]
head(final, 10)

#c: Top 10 Girl names for the client
FinalGirl <- subset(final[order(-final$Total),], final$Gender == "F")

#Remove extra columns
FinalGirl <- FinalGirl[,-4]
FinalGirl <- FinalGirl[,-3]
FinalGirl <- FinalGirl[,-2]

#d: Top 10 Girl names to a file
write.csv(file = "Top10GirlNames.csv", x = FinalGirl,row.names = F, quote = F)
```

###Resources

Input delimiters
http://www.sthda.com/english/wiki/reading-data-from-txt-csv-files-r-base-functions

Input help
https://www.statmethods.net/input/contents.html
