---
title: "6306-404 HW6 "
author: "Hayley Horn"
date: "October 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(plyr)
library(sqldf)
```

1. Mental Health Clinics

Load the file,and some basic clean-up
```{r Mental Health Clinics, ECHO=TRUE }
#a 
load("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_6/N-MHSS-2015-DS0001-data//N-MHSS-2015-DS0001-data-r.rda")
state <- data.frame(mh2015_puf$LST)
state2 <- data_frame(mh2015_puf$LST, mh2015_puf$FACILITYTYPE)
attach(state)
attach(state2)
#this throws off messages about masking "The following object is masked from state ..." that I don't know how to suppress


#clean data up enough to match
names(state) = "stateCode"
state$stateCode <- gsub(" ","",state$stateCode) #remove spaces
state$stateCode <- as.factor(state$stateCode)
names(state2) = c("stateCode","Type")

#join state with facilities list - I could not get this to work, wanted to join here and filter out VAMC
#state <- sqldf("SELECT state.*, state2.* FROM state INNER JOIN state2 USING(stateCode)")
#str(state)

```

Grab a distinct list of States
```{r distinct, ECHO=TRUE}
#b 
data.frame(state) %>% distinct()
```

Clean the file and remove off mainland states
```{r clean, ECHO = TRUE}
#this is repeated but the code will not work otherwise
state <- data.frame(state) #If I take this out, it will not work AGAIN
names(state) <- "stateCode"
attach(state)
#str(stateCode)

#c
rmstate <- data.frame(c("AK","HI","AS","GU","BR","VI"))
names(rmstate)  <- "stateCode"
#head(paste(stateCode),20) #Verify no leading or tailing spaces

#group by state code- simplifies my plot
pstate <- data.frame(sqldf("SELECT stateCode,COUNT(stateCode) as Count FROM state group by stateCode "))
#str(pstate)

#d Plot the amount by states
ggplot(pstate, aes(x = reorder(stateCode, -Count), y = (Count), fill = Count)) + geom_bar(stat = "identity", width = .75) + guides(fill = FALSE) + theme(legend.position = "none") + theme(plot.title = element_text(hjust = 0.5))+ ggtitle("VA Hospitals in the Continental US") + xlab("State") + ylab("# of Hospitals") + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))
```

2. Cleaning and Bringing in New Features (60%) 
 
Read in the size file
```{r ssize, ECHO = TRUE}
#a read in size file
ssize <- data.frame(read.csv("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_6/statesize.csv"))
names(ssize) <- c("StateName","SqMiles","stateCode","Region")
attach(ssize)
#If I don't add this part, 
state <- data.frame(state) #Is this is the 5th time I have loaded this? 
#str(pstate)
names(state) <- c("stateCode") #standardize name
#state$stateCode <- gsub(" ","",state$stateCode)
attach(state)
```

Make sure data is ready to merge and merge
```{r data prep, ECHO = TRUE}
#2a + 2b issue with leading and trailing spaces was identified on import and fixed in section one. I would like to develop a regex to run and look for standard dq issues, like leading trailing spaces, unprintable chars, nas and obscenities. The last option is standard for when I check data sources that either come from or get published to the web, and also if customer service employees have entered the data.
```


Add in Region information and plot
```{r plot, ECHO = TRUE}

#c join files

#d region and plot

reg <- read.csv("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")

names(reg) <- c("State","stateCode","Region","Division")

pstate <- sqldf("SELECT pstate.*, Region FROM pstate INNER JOIN reg USING(stateCode)")

df3 <- sqldf("SELECT distinct stateCode, SqMiles, pstate.*
              FROM ssize
             INNER JOIN pstate USING(stateCode)")
#str(df3)

p2state <- data.frame(df3, df3$SqMiles/1000)
names(p2state) <- c("stateCode","SqMiles","stateCode2","Count","Region","per1000SqMiles")
#str(p2state)

#Plot it all together
ggplot(p2state, aes(x = reorder(stateCode, - per1000SqMiles), y =(per1000SqMiles),  fill = Region)) + geom_bar(stat = "identity", width = .75) + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("VA Hospitals in the Continental US") + xlab("State") + ylab("# of Hospitals") + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
#color buckets should be not the default, but I couldn't make that work. I think it is using facets.

```


Summary
I am not good at r coding and it looks like my dog did this. He didn't.
Hadley Wickham would be very disappointed in this document.
