---
title: "6306-404 U2 live session Assignment"
author: "Hayley Horn"
date: "September 9, 2018"
output: html_document  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  
```  
####1a What is a basic workflow for reproducible research?
Data Gathering, Analysis and Presentation  

####1b What are five practical tips for making research reproducible?
1. Document everything. Note where you got the information, format so there is no ambiguity.

2. Everything is a text file. There is complexity in formatting, use a text file for simplicity.

3. All files should be human readable. Narrate your process inline with commenting and clear process.

4. Explicitly tie your files together. File relationships can be unclear, make it clear and don't leave room for error with unclear keys.

5. Have a plan to organize, store, and make your files available.   

####1c Give an example of how you might implement each tip.
1. There may be multiple data sources, if one works better than others, I can comment and note the decision process. It's not always obvious on revisit.

2. This is easy, after being tormented by an unprintable character bonking a real time integration, text file with a good text editor can save your hair.

3. To keep things readable, have another set of eyes review, and even better if it is someone that will tell me if they don't understand,

4. Joins need to be clear. Improper joins are a great way to bonk all my data at once. 

5. Establish my end of research plan up front, and use a process that I would hope someone would leave for my team. Data version of do unto others.  

####1d	Which one of these do you think will be the most difficult?
The last step of making the plan to store and share. If it is an externally held data source, there is a risk of the file being pulled or moved, so this is where, along with the first step that they should know what they are looking at.  

####2a From airquality data, plot temp vs month
```{r A2 Q2A, echo = TRUE}
plot(x=airquality$Month, y=airquality$Temp, xlab="Month", ylab="Temperature", main="Temperature by Month")
```

####2b From airquality data, plot temp vs Ozone
```{r A2 Q2B, echo = TRUE}
plot(x=airquality$Temp, y=airquality$Ozone, xlab="Temperature", ylab="Ozone", main="Temperature vs Ozone")
```

####3a From pressure data, plot temperature vs pressure
```{r pressure_plot, echo=TRUE}
plot(pressure$pressure,pressure$temperature, xlab="Pressure",ylab="Temperature", main="Temperature vs Pressure")
```

####3b From pressure data, plot presssure vs temperature
```{r pressure_plot2, echo=TRUE}
plot(pressure$temperature,pressure$pressure, xlab="Temperature",ylab="Pressure", main="Pressure vs Temperature")
```