---
title: "Codebook for 6306-404 HW5"
author: "Hayley Horn"
date: "October 1, 2018"
output: html_document
---

```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(pander)
```

###Purpose
The markdown file HayleyHorn_6306_404_U5.Rmd was used to practice data munging, data merging, and data summarizing.

###Summary choices
The customer wanted the top girls names of 2015 and 2016, so after cleanup and blending, the output was provided with limited fields and sorted descending.

###Raw Data 
Files: yob2016.txt; yob2015.txt


*yob2015.txt Codebook summary table*

-------------------------------------------------------------------
Label   Variable       Class       # unique  Missing  Description  
                                     values                        
------- -------------- --------- ---------- --------- -------------
        **[Name]**     factor         30553  0.00 %                

        **[Gender]**   factor             2  0.00 %                

        **[Count]**    integer         1397  0.00 %                
-------------------------------------------------------------------




---------------------------------
Variable list              Result
-------------------------------
Number of observations      33063
Number of variables             3
---------------------------------

-------------------------------------
Name                           Result
------------------------------------
Variable type                  factor
Number of missing obs.        0 (0 %)
Number of unique values         30553
Mode                        "Aalijah"
-------------------------------------

-----------------------------------
Gender                      Result
---------------------------------
Variable type                factor
Number of missing obs.      0 (0 %)
Number of unique values           2
Mode                            "F"
-----------------------------------

------------------------------------
Count                       Result
-----------------------------------
Variable type                integer
Number of missing obs.       0 (0 %)
Number of unique values         1397
Median                            11
1st and 3rd quartiles          7; 30
Min. and max.               5; 20415
------------------------------------
Source
Report generation information:
 *  Created by Hayley (username: `hayle`).
 *  Report creation time: Mon Oct 01 2018 19:15:44
 *  Report Was run from directory: `C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_3`
 *  dataMaid v1.1.2 [Pkg: 2018-05-03 from CRAN (R 3.5.1)]
 *  R version 3.5.1 (2018-07-02).
 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows >= 8 x64 (build 9200)).
 *  Function call: `makeDataReport(data = y2015, mode = "summarize", file = "codebook_y2015.Rmd", 
    checks = list(list("showAllFactorLevels")), listChecks = FALSE, 
    maxProbVals = FALSE, codebook = TRUE, reportTitle = "Codebook for y2015")`

*yob2016.txt Codebook summary table*
---------------------------------
Variable list                    Result
------------------------ --------
Number of observations      32868
Number of variables             3
---------------------------------

-----------------------------------
Name                        Result
------------------------------------
Variable type                  factor
Number of missing obs.        0 (0 %)
Number of unique values         30294
Mode                        "Aalijah"
-------------------------------------

-----------------------------------
Gender                      Result
----------------------------------
Variable type                factor
Number of missing obs.      0 (0 %)
Number of unique values           2
Mode                            "F"
-----------------------------------

------------------------------------
Count                       Result
-----------------------------------
Variable type                integer
Number of missing obs.       0 (0 %)
Number of unique values         1385
Median                            12
1st and 3rd quartiles          7; 30
Min. and max.               5; 19414
------------------------------------
Report generation information:
 *  Created by Hayley (username: `hayle`).
 *  Report creation time: Mon Oct 01 2018 19:15:42
 *  Report Was run from directory: `C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_3`
 *  dataMaid v1.1.2 [Pkg: 2018-05-03 from CRAN (R 3.5.1)]
 *  R version 3.5.1 (2018-07-02).
 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows >= 8 x64 (build 9200)).
 *  Function call: `makeDataReport(data = y2016, mode = "summarize", file = "codebook_y2016.Rmd", 
    checks = list(list("showAllFactorLevels")), listChecks = FALSE, 
    maxProbVals = FALSE, codebook = TRUE, reportTitle = "Codebook for y2016")`

###Tidy Data
*FinalGirl.csv Codebook summary table*
---------------------------------
Variable list                    Result
--------------------------------
Number of observations      15267
Number of variables             2
---------------------------------

Codebook summary table
------------------------------------------------------------------
Label   Variable      Class       # unique  Missing  Description  
                                    values                        
------------------------------------------------------------------
        **[Name]**    factor         15267  0.00 %                
        **[Total]**   integer         1300  0.00 %                
------------------------------------------------------------------

-------------------------------------
Name                        Result
----------------------------------
Variable type                 integer
Number of missing obs.        0 (0 %)
Number of unique values          1300
Median                             30
1st and 3rd quartiles          17; 79
Min. and max.               10; 39829
-------------------------------------
Source
Report generation information:
 *  Created by Hayley (username: `hayle`).
 *  Report creation time: Mon Oct 01 2018 23:04:47
 *  Report Was run from directory: `C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_3`
 *  dataMaid v1.1.2 [Pkg: 2018-05-03 from CRAN (R 3.5.1)]
 *  R version 3.5.1 (2018-07-02).
 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows >= 8 x64 (build 9200)).
 *  Function call: `makeDataReport(data = FinalGirl, mode = "summarize", file = "codebook_FinalGirl.Rmd", 
    checks = list(list("showAllFactorLevels")), listChecks = FALSE, 
    maxProbVals = FALSE, codebook = TRUE, reportTitle = "Codebook for FinalGirl")`


###Transformations

```{r `echo = FALSE`}
#Load the semicolon delimitied text file
y2016 <- data.frame(read.delim("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_5/yob2016.txt", header = FALSE, sep = ";", dec = "."))
#Give the columns names
names(y2016) <- c("Name","Gender","Count")
#Remove the duplicate row
y2016 <- y2016[-212,]
#Load the comma delimited text file
y2015 <- data.frame(read.delim("C:/Users/hayle/Dropbox/SMU/MSDS 6306/Unit_5/yob2015.txt", header = FALSE, sep = ",", dec = "."))
#Give the columns names
names(y2015) <- c("Name","Gender","Count")
#Merge
final <- (merge(y2016, y2015, by = c("Name","Gender")))
names(final) <- c("Name","Gender","2016_Count", "2015_Count")
```



